Globally-centered autocovariances in MCMC

### 2020-08-16 Paper/main.tex
Sunday morning changes

The power of the modern personal computer has made it easy to run parallel Markov chain Monte Carlo (MCMC) implementations. This is particularly useful for slow mixing or multi-modal target distributions, where the starting points of the chains are spread over the state-space in order to more accurately capture characteristics of the target distribution. Output from $m$ parallel chains are then summarized, visually and quantitatively, to assess the empirical mixing properties of the chains and the quality of Monte Carlo estimators.

@@

Let $F$ be the target distribution with mean $\mu$ defined on the set $\X \subseteq \mathbb{R}^p$ equipped with a countably generated Borel $\sigma$-field $\mathcal{B}(\X)$. For $s = 1, \dots, m$, let $\{X_{st}; \; t \in \mathbb{Z}\}$ be the $s^{th}$ $F$-Harris ergodic stationary Markov chain \citep[see][for definitions]{meyn:twee:2009} employed to learn characteristics about $F$.
% If $\bar{X} = n^{-1} \sum_{t=1}^{n}X_t$ , then with probability 1, $\bar{X} \to \mu$ as $n \to \infty$. \\
The process is covariance stationary so that
the ACvF at lag $k$ depends only on $k$ and is defined as 
%
\[
    \Gamma(k) = \Cov_F(X_{s1}, X_{s\,1+k})= \mathbb{E}_F \left[(X_{s1} - \mu)(X_{s\,1+k} - \mu)^T \right]\,.
\]
Estimating $\Gamma(k)$ is critical to assessing the quality of the sampler and the quality of sample statistics. Let $\bar{X}_s = n^{-1} \sum_{t=1}^{n} X_{st}$ denote the Monte Carlo estimator of $\mu$ from the $s$th chain. The standard estimator is the sample autocovariance matrix at lag $k$ for the $s$th chain is
%
\begin{equation} \label{eq:empirical_acf}
    \hat{\Gamma}_s(k) = \dfrac{1}{n}\sum_{t=1}^{n-\abs{k}} \left(X_{st} - \bar{X}_s \right) \left(X_{s t+\abs{k}} - \bar{X}_s \right)^T\,.
\end{equation}
%
% It is a known fact that most of the Markov chains obtained from practical MCMC algorithms are not stationary. Regardless, Equation~\ref{eq:empirical_acf} provides the empirical estimator for ACF at a particular lag for the Markov chain with any initial distribution (\cite{geyer2011introduction}).\\

% Calculation of exact ACF has been done for simple stationary stochastic models like autoregressive (AR) models (under assumption of stationarity), moving average (MA) (\cite{quenouille1947notes}), and autoregressive-moving-average (ARMA) (\cite{box2015time}) models. However, for most of the complex sampling algorithms (including MCMC), we rely on sample estimators for ACF. Unfortunately, the estimator given in Equation~\ref{eq:empirical_acf} has been observed to have a some drawbacks. Firstly, it is not immune to outliers (see \cite{ma2000highly}) and secondly, it is usually non-informative in case of slowly mixing Markov chains. 
For single-chain MCMC runs, the estimator $\hat{\Gamma}(k)$ is used to construct autocorrelation (ACF) plots, to estimate the long-run variance of Monte Carlo estimators \cite{hannan:1970,dame:1991}, and to estimate effective sample size for an estimation problem \cite{kass:carlin:gelman:neal:1998,gong:fleg:2016,vats:fleg:jon:2019}.

@@

 % A commonly faced problem with the MCMC models is the slow mixing property. Especially in case of multi modal target distributions, it has been observed that the Markov chains tend to get \textit{stuck} in a particular mode for a long time. In such situations, single chain empirical ACF will severely underestimate the autocovariance. 
 We propose a globally-centered ACvF (G-ACvF) that centers the Markov chain around the overall mean vector from all $m$ chains. We show that the bias under stationarity for G-ACvF is lower than $\hat{\Gamma}(k)$ and through various examples empirically demonstrate improved estimation. We employ the G-ACvF estimators to construct autocorrelation plots which leads to remarkably improvements. A demonstrative example is at the end of this Section.

@@

ACvFs are required to estimate the log-run variance of Monte Carlo averages. Specifically, spectral variance estimators are used to estimate $\Sigma = \sum_{k=-\infty}^{\infty} \Gamma(k)$. We replace $\hat{\Gamma}(k)$ with G-ACvF in spectral variance estimators and show strong consistency of this estimator under weak conditions. In the spirit of \cite{andr:1991}, we also obtain large-sample bias and variance of the resulting estimator of $\Sigma$. Spectral variance estimators can be slow to prohibitively expensive to calculate \citep{liu:fleg:2018}. We relieve the computational burden, we adapt the fast algorithm of \cite{heberle2017fast} for G-ACvFs that dramatically reduces computation time. The globally-centered spectral variance estimators in employed in the computation of effective sample size and safeguards against early termination of the MCMC process and improved reliability of estimators.

@@

We demonstrate the striking difference in estimation of $\Gamma(k)$ via ACF plots. Consider a random-walk Metropolis-Hastings sampler for a univariate mixture of Gaussians target density. That is, $F$ is the distribution with density
\[
f(x) = 0.7\,f(x; -5, 1) + 0.3\,f(x; 5, 0.5)\,,
\]
where $f(x; a,b)$ is the density of a normal distribution with mean $a$ and variance $b$. 
Two Markov chains are started at each mode.  Figure~\ref{fig:gaussian-trace} indicates that in the first $10^4$ samples, the chains do not jump modes so that both Markov chains yield different estimates of the population mean, $\mu$. At $10^5$ sample size, both Markov chains have traversed the state space reasonably and yield similar estimates of $\mu$. We make two points here: first, for small sample size, we observe that the locally-centered ACF gives misleading estimates   whereas G-ACF accounts for the discrepancy in sample means. This is evident from the ACF plots  in Figure~\ref{subfig:gauss-acf-1e4}. Second, for a large sample size, the estimates from G-ACF as well as ACF are equivalent in Figure \ref{subfig:gauss-trace_1e5}.

@@

\begin{figure}[htbp]
\begin{subfigure}{\textwidth}
   \centering
   \includegraphics[width=.9\linewidth]{plots/gaussian-acf_n1e4.pdf}
   \caption{$n = 10^4$}
   \label{subfig:gauss-acf-1e4}
 \end{subfigure}\\
 \begin{subfigure}{\textwidth}
   \centering
   \includegraphics[width=.9\linewidth]{plots/gaussian-acf_n1e5.pdf} 
   \caption{$n = 10^5$}
   \label{subfig:gauss-acf-1e5}
 \end{subfigure}
 \caption{Autocorrelation plot for locally-centered ACF (left) and globally centered ACF (right) averaged over all chains. (a) $n = 10^4$, not converged yet; (b) $n =  10^5$, converged}
 \label{fig:gauss:acf}
\end{figure}

@@

Recall that we run $m$ independent $F$-ergodic Markov chains denoted by $\{X_{st}; \; t \in \mathbb{Z}\}$ for the $s^{th}$ chain. Let the sample mean of the $s^{th}$ Markov chain be $\bar{X}_s = n^{-1} \sum_{t=1}^{n} X_{st}$ and the global mean by $\bar{\bar{X}} = m^{-1}\sum_{s = 1}^{m}\bar{X}_s$. The global mean is naturally superior estimator of $\mu$ than $\bar{X}_s$. We define the globally-centered autocovariance function (G-ACvF) estimator for the $s^{th}$ Markov chain as
%
\[
\hat{\Gamma}_{G,s}(k) = \dfrac{1}{n} \sum_{t=1}^{n-k}(X_{st}-\bar{\bar{X}})(X_{s(t+k)}-\bar{\bar{X}})^T,
\]
In the event that all $m$ Markov chains have been run long enough that $\bar{X}_s \approx \bar{\bar{X}}$ are similar, then  $\hat{\Gamma}_{s} (k) \approx \hat{\Gamma}_{G,s}(k)$. However, for shorter runs or for slow mixing chains, $\Gamma(k)$ is more appropriately estimated by $\hat{\Gamma}_{G,s}$ as it utilizes information from all chains and accounts for disparity between estimates of $\mu$. We quantify the gains in  bias  for the G-ACvF estimator below.
% If the starting points of these parallel Markov chains are sufficiently dispersed, G-ACF provides more realistic estimation of lag covariances. \cite{priestley1981spectral} shows that the empirical estimator in Equation \ref{eq:empirical_acf} is biased with the main bias term equal to $\abs{k}\Gamma(k)/n$, ignoring the small order terms arising due to estimation of $\mu$. Though, an unbiased estimator with a divisor of $n - \abs{k}$ instead of $n$ exists, the biased estimator is preferred for its lower mean square error for larger relative $k$  (\cite{priestley1981spectral})
Let 
\[
\Phi^{(q)} = \sum_{ k= -\infty}^{\infty}\abs{k}^q \Gamma(k)\,,
\]
and let $\Phi^{(1)}$ be denoted by $\Phi$.  The proof of the following theorem is available in Appendix~\ref{appendix:bias}.

\begin{theorem} \label{th:G-_expec} Let $\E_F \|X_{11}\|^{2 + \delta} < \infty$ for some $\delta > 0$. Let $P$ be a  polynomially ergodic Markov chain of order $\xi > (2 + \epsilon)/(1 + 2/\delta)$ for some $\epsilon > 0$. Then,
\[
   \mathbb{E}_F\left[\hat{\Gamma}_{G,s}(k) \right] = \left(1- \dfrac{|k|}{n}\right) \left(\Gamma(k) - \dfrac{\Sigma}{mn} - \dfrac{\Phi}{mn^2}\right)  + o \left(n^{-2} \right)\,,
\]
and consequently,
\[
\mathbb{E}_F\left[\hat{\Gamma}_{G,s}(k) - \hat{\Gamma}_s(k) \right] = -\dfrac{\abs{k}}{n}\Gamma(k) - \dfrac{1}{m}\left(1- \dfrac{|k|}{n}\right)\left(\dfrac{\Sigma}{n} + \dfrac{\Phi}{n^2}\right) + o\left(n^{-2} \right)\,. 
\]
\end{theorem}

@@

When $m = 1$, $\hat{\Gamma}_{G,s}(k) = \hat{\Gamma}_{s}(k)
$ the bias results for which can be found in \cite{priestley1981spectral}. Since lag-covariances are typically positive in MCMC, Theorem~\ref{th:G-_expec} implies that the G-ACvF estimators are asymptotically unbiased and exhibit reduced bias in finite samples compared to locally-centered ACvF estimators. The consequences of this are particularly pertinent for the diagonals of $\Gamma(k)$.
 % Special interest is also in the lag-0 autocovariance, that is, the variance of the target distribution. 
\begin{remark} \label{rem:lag0_expectation}
The variance of the target distribution is the lag covariance at lag 0 and is of particular interest. The bias of the estimator reduces by a factor of $m$:
\[
\mathbb{E} \left[\hat{\Gamma}_{G}(0) \right] = \Gamma(0) - \dfrac{\Sigma}{mn} - \dfrac{\Phi}{mn^2} + o(n^{-2})
\]
\end{remark}

A first use of the ACvFs in MCMC is in constructing ACF plots. Let the target variance of $i${th} component be $\Gamma^{ii}(0)$. For any component $i$, the autocorrelation is defined as
\[
\rho^{ii}(k) = \dfrac{\Gamma^{ii}(k)}{\Gamma^{ii}(0)}\,,
\]
and is instrumental in visualizing the serial correlation in components of the Markov chain. The typical estimate of the autocorrelation is constructed from $\hat{\Gamma}_s(k)$. Instead, we advocate for using G-ACvF estimates so that,
\[
\hat{\rho}_{G,s}^{(i)}(k) = \dfrac{ \hat{\Gamma}^{(i)}_{G,s} (k)}{\hat{\Gamma}^{(i)}_{G,s} (0)}.
\]
The globally-centered autocorrelation provides a far more realistic assessment of the correlation structure of marginal components of the chain  as evidenced in Figure~\ref{fig:gauss:acf}.

@@

Set $\{Y_{st}\}_{t \geq 1} = \{g(X_{st})\}_{t \geq 1}$ for all $s = 1, \dots, m$ and set $\bar{Y}_s = n^{-1}\sum_{t=1}^{n}Y_{st}$  and  $\bar{\bar{Y}} = m^{-1}\sum_{s=1}^{m}\bar{Y}_s$. By Birkhoff's ergodic theorem,  $\bar{\bar{Y}} \to \mu_g$ as $n \to \infty$ with probability 1.

Reliability of $\bar{\bar{Y}}$ is assessed by its Monte Carlo variability via an asymptotic sampling distribution \citep{fleg:hara:jone:2008,roy:2019,vats:rob:fle:jon:2020}. 
% To assess the reliability of simulations, we are interested in Monte Carlo error, i.e. $\bar{Y_s} - \mu$ for the $s^{th}$ chain. 
% The sampling distribution for Monte Carlo error is available via Markov chain CLT (\cite{jones2004markov})
 A Markov chain central limit theorem (CLT) holds if there exists a $p \times p$ positive-definite matrix $\Sigma$ such that
%
\begin{equation}
\label{eq:CLT}
  \sqrt{mn}(\bar{\bar{Y}} - \mu) \xrightarrow{d} N(0,\Sigma)\,
\end{equation}
where
\begin{equation}
\label{eq:sigma}
  \Sigma = \sum_{k = -\infty}^{\infty} \Cov_F \left( Y_{11}, Y_{1(1+k)} \right)
\end{equation}
%
The goal in output analysis for MCMC is to estimate $\Sigma$in order to assess variability in $\bar{\bar{Y}}$. There is a rich literature of estimators of $\Sigma$ available for single-chain MCMC implementations. The most common are spectral variance estimators (SVE) \citep{andr:1991,vats:fleg:jon:2018} and batch means estimators \cite{chen:seila:1987,vats:fleg:jon:2019}. Recently, \cite{gupta:vats:2020} constructed a replicated batch means estimator for estimating $\Sigma$ from multiple parallel chains. Batch means estimators are computationally more efficient than SVE, SVE estimators are more reliable \citep{fleg:jone:2010}. Here, we utilize G-ACvF estimators to construct globally-centered SV estimator (G-SVE) of $\Sigma$. Using the methods of \cite{heberle2017fast}, we also provide an efficient algorithm for obtaining the G-SV estimator.

SV estimators are constructed as weighted and truncated sums of estimated lag covariances. Let $w: \mathbb{R} \to [-z,z]$ be a lag window function and let $b_n \in \mathbb{N}$ be a truncation point. The SV estimator of $\Sigma$ from a single chain is
%
{\color{blue} DV: limits should be $-n +1$ and $n - 1$.}
\begin{equation} \label{eq:sve}
    \hat{\Sigma}_{SV,s} = \sum_{k=-b_n+1}^{b_n-1}w\left(\dfrac{k}{b_n}\right)\hat{\Gamma}_s(k)
\end{equation}
An overall estimate of $\Sigma$ from all $m$ chains is the average SV estimator (ASV):
\[
\hat{\Sigma}_A = \dfrac{1}{m}\sum_{s=1}^{m}\hat{\Sigma}_{SV,s}\;.
\]
%  We propose a globally-centered spectral variance estimator (G-SVE) wherein we use G-ACF to estimate the autocovariance. Strong consistency for MSVE has been addressed by \cite{vats:fleg:jon:2018}. Bias and variance calculations for a variant of SVE with known mean is done by \cite{hannan2009multiple}. In this paper, we will address the asymptotic properties (strong consistency), bias and variance calculations for G-SVE.

@@

% Recall from Equation \ref{eq:empirical_acf} that $\hat{\Gamma}(k)$ is the empirical autocovariance function given by

% \begin{equation}
%     \hat{\Gamma}(k) = \dfrac{1}{n}\sum_{t=1}^{n-\abs{k}}(X_t - \bar{X})(X_{t + \abs{k}} - \bar{X})^T
% \end{equation}
{\color{blue} DV: There are some issues.
\begin{itemize}
  \item $|w_n| < z$ or $|w_n| < c$?
  \item Need new notation for $\Gamma(k)$ since $\Gamma(k)$ is for the $X$ and we need a new notation for $Y$. This change needs to be made for this whole section and the appendix!
\end{itemize}
}

We define the globally-centered SV estimator as the weighted and truncated sum of G-ACvFs
%
% Consider $m$ Markov chains where $\hat{\Gamma}_s(k)$, $\hat{\Gamma}_{G,s}(k)$ and $\hat{\Sigma}_{S, s}$ are the lag-$k$ empirical ACF, G-ACF, and SVE respectively for $s^{th}$ Markov chain, $s\in {1,...,m}$. 
% Consider the globally-centered SV estimator (G-SVE) for the $s$th chain using G-ACF:
%
\begin{align*}
    \hat{\Sigma}_{G} &= \sum_{k=-b_n+1}^{b_n-1}w\left(\dfrac{k}{b_n}\right)\hat{\Gamma}_{G}(k)\,.
\end{align*}
% and the combined G-SVE is
% \[
% \hat{\Sigma}_{G} =  \dfrac{1}{m}\sum_{s=1}^{m}\hat{\Sigma}_{G,s}\,.
% \]
For $\hat{\Sigma}_{G}$ to accurately estimate $\Sigma$, conditions need to be imposed on the weight function, $w_n$.

{\color{blue}DV: Did we discuss that $W_n$ should be define as an infinite sum and should be $W$?}

>>>>

The power of the modern personal computer has made it easy to run parallel Markov chain Monte Carlo (MCMC) implementations. This is particularly useful for slow mixing or multi-modal target distributions, where the starting points of the chains are spread over the state-space in order to capture characteristics of the target distribution more accurately. The output from $m$ parallel chains is then summarized, visually and quantitatively, to assess the empirical mixing properties of the chains and the quality of Monte Carlo estimators.

@@

Let $F$ be the target distribution with mean $\mu$ defined on the set $\X \subseteq \mathbb{R}^p$ equipped with a countably generated Borel $\sigma$-field $\mathcal{B}(\X)$. For $s = 1, \dots, m$, let $\{X_{st}; \; t \in \mathbb{Z}\}$ be the $s^{th}$ $F$-Harris ergodic stationary Markov chain \citep[see][for definitions]{meyn:twee:2009} employed to learn characteristics about $F$.
% If $\bar{X} = n^{-1} \sum_{t=1}^{n}X_t$ , then with probability 1, $\bar{X} \to \mu_g$ as $n \to \infty$. \\
The process is covariance stationary so that
the ACvF at lag $k$ depends only on $k$ and is defined as 
%
\[
    \Upsilon(k) = \Cov_F(X_{s1}, X_{s\,1+k})= \mathbb{E}_F \left[(X_{s1} - \mu)(X_{s\,1+k} - \mu)^T \right]\,.
\]
Estimating $\Gamma(k)$ is critical to assessing the quality of the sampler and the quality of sample statistics. Let $\bar{X}_s = n^{-1} \sum_{t=1}^{n} X_{st}$ denote the Monte Carlo estimator of $\mu$ from the $s$th chain. The standard estimator for $\Gamma(k)$ is the sample autocovariance matrix at lag $k$. For the $s$th chain, it is given by
%
\begin{equation} \label{eq:empirical_ACvF}
    \hat{\Gamma}_s(k) = \dfrac{1}{n}\sum_{t=1}^{n-\abs{k}} \left(X_{st} - \bar{X}_s \right) \left(X_{s\, t+\abs{k}} - \bar{X}_s \right)^T\,.
\end{equation}
%
% It is a known fact that most of the Markov chains obtained from practical MCMC algorithms are not stationary. Regardless, Equation~\ref{eq:empirical_ACvF} provides the empirical estimator for ACF at a particular lag for the Markov chain with any initial distribution (\cite{geyer2011introduction}).\\

% Calculation of exact ACF has been done for simple stationary stochastic models like autoregressive (AR) models (under assumption of stationarity), moving average (MA) (\cite{quenouille1947notes}), and autoregressive-moving-average (ARMA) (\cite{box2015time}) models. However, for most of the complex sampling algorithms (including MCMC), we rely on sample estimators for ACF. Unfortunately, the estimator given in Equation~\ref{eq:empirical_ACvF} has been observed to have a some drawbacks. Firstly, it is not immune to outliers (see \cite{ma2000highly}) and secondly, it is usually non-informative in case of slowly mixing Markov chains. 
For single-chain MCMC runs, the estimator $\hat{\Gamma}(k)$ is used to construct autocorrelation (ACF) plots, to estimate the long-run variance of Monte Carlo estimators \cite{hannan:1970,dame:1991}, and to estimate effective sample size for an estimation problem \cite{kass:carlin:gelman:neal:1998,gong:fleg:2016,vats:fleg:jon:2019}.

@@

 % A commonly faced problem with the MCMC models is the slow mixing property. Especially in case of multi modal target distributions, it has been observed that the Markov chains tend to get \textit{stuck} in a particular mode for a long time. In such situations, single chain empirical ACF will severely underestimate the autocovariance. 
 We propose a globally-centered ACvF (G-ACvF) that centers the Markov chain around the overall mean vector from all $m$ chains. We show that the bias under stationarity for G-ACvF is lower than $\hat{\Gamma}(k)$, and through various examples, empirically demonstrate improved estimation. We employ the G-ACvF estimators to construct autocorrelation plots, which leads to remarkable improvements. A demonstrative example is at the end of this Section.

@@

ACvFs are required to estimate the long-run variance of Monte Carlo averages. Specifically, spectral variance estimators are used to estimate $\Sigma = \sum_{k=-\infty}^{\infty} \Gamma(k)$. We replace $\hat{\Gamma}(k)$ with G-ACvF in spectral variance estimators (SVE) and show strong consistency of this estimator under weak conditions. In the spirit of \cite{andr:1991}, we also obtain large-sample bias and variance of the resulting estimator of $\Sigma$. Spectral variance estimators can be slow to prohibitively expensive to calculate \citep{liu:fleg:2018}. To relieve the computational burden, we adapt the fast algorithm of \cite{heberle2017fast} for G-ACvFs that dramatically reduces computation time. The globally-centered spectral variance estimator (G-SVE) is employed in the computation of effective sample size through the estimator given by \cite{vats2019multivariate}. We will show that using G-SVE for estimating $\Sigma$ safeguards us against early termination of the MCMC process and improves the reliability of estimators.

@@

We demonstrate the striking difference in estimation of $\Gamma(k)$ via autocorrelation function (ACrF) plots. Consider a random-walk Metropolis-Hastings sampler for a univariate mixture of Gaussians target density. Let the target density be 
\[
f(x) = 0.7\,f(x; -5, 1) + 0.3\,f(x; 5, 0.5)\,,
\]
where $f(x; a,b)$ is the density of a normal distribution with mean $a$ and variance $b$. 
Two Markov chains are started at each mode. Let $n$ denote the number of samples for each chain. Figure~\ref{fig:gaussian-trace} indicates that in the first $10^4$ samples, the chains do not jump modes so that both Markov chains yield different estimates of the population mean, $\mu$. At $10^5$ sample size, both Markov chains have traversed the state space reasonably and yield similar estimates of $\mu$. We make two points here: first, for small sample size, we observe that the locally-centered ACF gives misleading estimates whereas G-ACF accounts for the discrepancy in sample means between two chains. This is evident from the ACF plots  in Figure~\ref{subfig:gauss-acf}. Second, for a large sample size, the estimates from G-ACrF as well as ACrF are equivalent. Therefore, G-ACF can be used in any situation with a promise of estimates that are at least as good as empirical ACF estimators.

@@

\begin{figure}[htbp]
\centering
\begin{subfigure}{.9\textwidth}
   \centering
   \includegraphics[width=.9\textwidth]{plots/gaussian-acf_hist.pdf}
   \caption{$n-10^4$ (histogram) and $n = 5 \times 10^5$ (blue line)}
   \label{subfig:gauss-acf_1e4_1e5}
\end{subfigure}
\begin{subfigure}{.9\textwidth}
   \centering
   \includegraphics[width=.9\textwidth]{plots/gaussian-acf_1e4.pdf} 
   \caption{$n = 10^4$}
   \label{subfig:gauss-acf_1e4}
\end{subfigure}
\caption{ACF plots for locally-centered ACF (left) and globally centered ACF (right). (a) The histogram plots the estimates for $n = 10^4$ and the blue line plot the estimates for $n = 5 \times 10^4$ for chain-1. (b) $n = 10^4$; the dotted lines are ACF for each chain and the thick solid line is the ACF averaged over both chains.}
\label{fig:gaussian-acf}
\end{figure}

@@

Recall that we run $m$ independent $F$-ergodic Markov chains denoted by $\{X_{st}; \; t \in \mathbb{Z}\}$ for the $s^{th}$ chain. Let the sample mean of the $s^{th}$ Markov chain be $\bar{X}_s = n^{-1} \sum_{t=1}^{n} X_{st}$ and the global mean by $\bar{\bar{X}} = m^{-1}\sum_{s = 1}^{m}\bar{X}_s$. The global mean is naturally superior estimator of $\mu$ than $\bar{X}_s$. We define the globally-centered autocovariance function (G-ACvF) estimator for the $s^{th}$ Markov chain as
%
\begin{equation} \label{eq:G-ACvF}
    \hat{\Gamma}_{G,s}(k) = \dfrac{1}{n} \sum_{t=1}^{n-k}(X_{st}-\bar{\bar{X}})(X_{s(t+k)}-\bar{\bar{X}})^T,
\end{equation}

In the event that all $m$ Markov chains have been run long enough that $\bar{X}_s \approx \bar{\bar{X}}$ are similar, then  $\hat{\Gamma}_{s} (k) \approx \hat{\Gamma}_{G,s}(k)$. However, for shorter runs or for slow mixing chains, $\Gamma(k)$ is more appropriately estimated by $\hat{\Gamma}_{G,s}$ as it utilizes information from all chains and accounts for disparity between estimates of $\mu$. We quantify the gains in  bias  for the G-ACvF estimator below.
% If the starting points of these parallel Markov chains are sufficiently dispersed, G-ACF provides more realistic estimation of lag covariances. \cite{priestley1981spectral} shows that the empirical estimator in Equation \ref{eq:empirical_ACvF} is biased with the main bias term equal to $\abs{k}\Gamma(k)/n$, ignoring the small order terms arising due to estimation of $\mu$. Though, an unbiased estimator with a divisor of $n - \abs{k}$ instead of $n$ exists, the biased estimator is preferred for its lower mean square error for larger relative $k$  (\cite{priestley1981spectral})
Let 
\[
\Phi^{(q)} = \sum_{ k= -\infty}^{\infty}\abs{k}^q \Gamma(k)\,,
\]
and let $\Phi^{(1)}$ be denoted by $\Phi$.  The proof of the following theorem is available in Appendix~\ref{appendix:bias}.

\begin{theorem} \label{th:G-ACF_bias} Let $\E_F \|X_{11}\|^{2 + \delta} < \infty$ for some $\delta > 0$. Let $P$ be a  polynomially ergodic Markov chain of order $\xi > (2 + \epsilon)/(1 + 2/\delta)$ for some $\epsilon > 0$. Then,
\[
   \mathbb{E}_F\left[\hat{\Gamma}_{G,s}(k) \right] = \left(1- \dfrac{|k|}{n}\right) \left(\Gamma(k) - \dfrac{\Sigma}{mn} - \dfrac{\Phi}{mn^2}\right)  + o \left(n^{-2} \right)\,,
\]
and consequently,
\[
\mathbb{E}_F\left[\hat{\Gamma}_{G,s}(k) - \hat{\Gamma}_s(k) \right] =  \left(1- \dfrac{|k|}{n}\right)\left(1- \dfrac{1}{m}\right)\left(\dfrac{\Sigma}{n} + \dfrac{\Phi}{n^2}\right) + o\left(n^{-2} \right)\,. 
\]
\end{theorem}

{\color{orange}MA: I corrected the RHS for the abover equation. Or did you mean to write $\mathbb{E}_F\left[\hat{\Gamma}_{G,s}(k) - \Gamma(k) \right]as LHS?$}

@@

When $m = 1$, $\hat{\Gamma}_{G,s}(k) = \hat{\Gamma}_{s}(k)
$ the bias results for which can be found in \cite{priestley1981spectral}. Since lag-covariances are typically positive in MCMC, Theorem~\ref{th:G-ACF_bias} implies that the G-ACvF estimators are asymptotically unbiased and exhibit reduced bias in finite samples compared to locally-centered ACvF estimators. The consequences of this are particularly pertinent for the diagonals of $\Gamma(k)$.
 % Special interest is also in the lag-0 autocovariance, that is, the variance of the target distribution. 
\begin{remark} \label{rem:lag0_expectation}
The variance of the target distribution is the lag covariance at lag 0 and is of particular interest. The bias of the estimator reduces by a factor of $m$:
\[
\mathbb{E} \left[\hat{\Gamma}_{G}(0) \right] = \Gamma(0) - \dfrac{\Sigma}{mn} - \dfrac{\Phi}{mn^2} + o(n^{-2})
\]
\end{remark}

A first use of the ACvFs in MCMC is in constructing ACF plots. Let the target variance of $i${th} component be $\Gamma^{ii}(0)$. For any component $i$, the autocorrelation is defined as
\[
\rho^{ii}(k) = \dfrac{\Gamma^{ii}(k)}{\Gamma^{ii}(0)}\,,
\]
and is instrumental in visualizing the serial correlation in components of the Markov chain. The typical estimate of the autocorrelation is constructed from $\hat{\Gamma}_s(k)$. Instead, we advocate for using G-ACvF estimates so that,
\[
\hat{\rho}_{G,s}^{(i)}(k) = \dfrac{ \hat{\Gamma}^{(i)}_{G,s} (k)}{\hat{\Gamma}^{(i)}_{G,s} (0)}.
\]
The globally-centered autocorrelation provides a far more realistic assessment of the correlation structure of marginal components of the chain  as evidenced in Figure~\ref{fig:gaussian:acf}.

@@

Set $\{Y_{st}\}_{t \geq 1} = \{g(X_{st})\}_{t \geq 1}$ for all $s = 1, \dots, m$. Let $\bar{Y}_s = n^{-1}\sum_{t=1}^{n}Y_{st}$  and  $\bar{\bar{Y}} = m^{-1}\sum_{s=1}^{m}\bar{Y}_s$. By Birkhoff's ergodic theorem,  $\bar{\bar{Y}} \to \mu_g$ as $n \to \infty$ with probability 1.

Reliability of $\bar{\bar{Y}}$ is assessed by its Monte Carlo variability via an asymptotic sampling distribution \citep{fleg:hara:jone:2008,roy:2019,vats:rob:fle:jon:2020}. 
% To assess the reliability of simulations, we are interested in Monte Carlo error, i.e. $\bar{Y_s} - \mu$ for the $s^{th}$ chain. 
% The sampling distribution for Monte Carlo error is available via Markov chain CLT (\cite{jones2004markov})
 A Markov chain central limit theorem (CLT) holds if there exists a $p \times p$ positive-definite matrix $\Sigma$ such that
%
\begin{equation}
\label{eq:CLT}
  \sqrt{mn}(\bar{\bar{Y}} - \mu_g) \xrightarrow{d} N(0,\Sigma)\,
\end{equation}
where
\begin{equation}
\label{eq:sigma}
  \Sigma = \sum_{k = -\infty}^{\infty} \Cov_F \left( Y_{11}, Y_{1(1+k)} \right) := \sum_{k = -\infty}^{\infty}\Upsilon (k)
\end{equation}
%
The goal in output analysis for MCMC is to estimate $\Sigma$ in order to assess variability in $\bar{\bar{Y}}$. There is a rich literature on estimators of $\Sigma$ available for single-chain MCMC implementations. The most common are spectral variance estimators (SVE) \citep{andr:1991,vats:fleg:jon:2018} and batch means estimators \citep{chen:seila:1987,vats:fleg:jon:2019}. Recently, \cite{gupta:vats:2020} constructed a replicated batch means estimator for estimating $\Sigma$ from multiple parallel chains. Batch means estimators are computationally more efficient than SVE, whereas SVE estimators are more reliable \citep{fleg:jone:2010}. Here, we utilize G-ACvF estimators to construct globally-centered SV estimator (G-SVE) of $\Sigma$. Using the methods of \cite{heberle2017fast}, we also provide an efficient algorithm for obtaining the G-SV estimator.

SV estimators are constructed as weighted and truncated sums of estimated lag covariances. The locally and globally centered empirical estimators for $\Upsilon(k)$, $\hat{\Upsilon}_s(k)$ and $\hat{\Upsilon}_{G,s}(k)$, can be defined in a similar way as $\hat{\Gamma}_s$ and $\hat{\Gamma}_{G,s}$ in Equation~\ref{eq:empirical_ACvF},\ref{eq:G-ACvF} for the Markov chain $\{Y_{st}\}_{t \geq 1}$ for all $s$. Let $w: \mathbb{R} \to [-c,c]$ be a lag window function and $b_n \in \mathbb{N}$ be a truncation point. The SV estimator of $\Sigma$ from a single chain is
%

\begin{equation} \label{eq:sve}
    \hat{\Sigma}_{s} = \sum_{k=-n+1}^{n-1}w\left(\dfrac{k}{b_n}\right)\hat{\Upsilon}_s(k)
\end{equation}

An overall estimate of $\Sigma$ from all $m$ chains is the average SV estimator (ASV):
\[
\hat{\Sigma}_A = \dfrac{1}{m}\sum_{s=1}^{m}\hat{\Sigma}_{s}\;.
\]
%  We propose a globally-centered spectral variance estimator (G-SVE) wherein we use G-ACF to estimate the autocovariance. Strong consistency for MSVE has been addressed by \cite{vats:fleg:jon:2018}. Bias and variance calculations for a variant of SVE with known mean is done by \cite{hannan2009multiple}. In this paper, we will address the asymptotic properties (strong consistency), bias and variance calculations for G-SVE.

@@

% Recall from Equation \ref{eq:empirical_ACvF} that $\hat{\Gamma}(k)$ is the empirical autocovariance function given by

% \begin{equation}
%     \hat{\Gamma}(k) = \dfrac{1}{n}\sum_{t=1}^{n-\abs{k}}(X_t - \bar{X})(X_{t + \abs{k}} - \bar{X})^T
% \end{equation}

We define the globally-centered SV estimator as the weighted and truncated sum of G-ACvFs
%
% Consider $m$ Markov chains where $\hat{\Gamma}_s(k)$, $\hat{\Gamma}_{G,s}(k)$ and $\hat{\Sigma}_{S, s}$ are the lag-$k$ empirical ACF, G-ACF, and SVE respectively for $s^{th}$ Markov chain, $s\in {1,...,m}$. 
% Consider the globally-centered SV estimator (G-SVE) for the $s$th chain using G-ACF:
%
\begin{align*}
    \hat{\Sigma}_{G} &= \sum_{k=-b_n+1}^{b_n-1}w\left(\dfrac{k}{b_n}\right)\hat{\Upsilon}_{G}(k)\,.
\end{align*}
% and the combined G-SVE is
% \[
% \hat{\Sigma}_{G} =  \dfrac{1}{m}\sum_{s=1}^{m}\hat{\Sigma}_{G,s}\,.
% \]
For $\hat{\Sigma}_{G}$ to accurately estimate $\Sigma$, conditions need to be imposed on the weight function, $w$. 
%

\begin{ass}
\label{ass:lag_window}
The lag window $w(x)$ is a continuous at $0$ and at all but finite number of points. It is bounded and even function with $w(0)=1, \; \abs{w(x)} < c$, \; $\int_{-\infty}^{\infty}w^2(x)dx < \infty$, and $W = \sum_{k = -\infty}^{\infty} w(k/b_n)$ is finite.
\end{ass}
% We will consider a range of lag windows $w(x)$; where $w(x)$ is a continuous, even function with $w(0)=1, \abs{w(x)} < c$, $\int_{-\infty}^{\infty}w^2(x)dx < \infty$, and $W_n < \infty$. 
Assumption~\ref{ass:lag_window} is standard \citep[see]{ande:1971} and is satisfied by most lag windows. We will use the popular Bartlett lag window in our simulations. The function $w(.)$ is given by
\[
w(x) = \begin{cases}
\left(1 - \abs{x}\right) & \textrm{for } \abs{x} \leq 1 \\
0 & \textrm{otherwise}\,.
\end{cases}
\]

In relation to lag window, we are interested in $w(k/b_n)$ for lag-$k$. Therefore, zero weightage is given to autocovariances for lag greater than the truncation point $b_n$. This is true for Bartlett but not true for the quadratic spectral lag window in \cite{andr:1991} for example.
